\ProjectName{}: Data Imputation as a Query Optimization

### 2017-02-23 paper/algorithm.tex
paper: another partial pass over algorithm section

\begin{align*}
  \textsc{Losses}(q) &= \begin{cases}
     1 :: \textsc{Losses}(q') & q = \delta_C(q') \\
    \frac{1}{\sqrt{ \textsc{Attr}(q') * \textsc{Card}(q') }}  :: \textsc{Losses}(q')  & q = \mu_C(q') \\
    \textsc{Losses}(q_1') ::: \textsc{Losses}(q_2') & q = q_1' \bowtie_\psi q_2' \\
    \textsc{Losses}(q') & q = \sigma_\phi(q'), \\ & q = \pi_C(q') \\
    Nil & q = t\\
  \end{cases}\\
  \textsc{Loss}(q) &= \frac{\Sigma_{l \in \textsc{Losses}(q)} l}{| \textsc{Losses}(q)|}
\end{align*}

@@

The planner operates as follows.
First, it collects the set of attributes which are used by the operators in the plan or which are visible in the output of the plan.
This set will be used to determine which attributes are imputed.
Then, it constructs a plan cache. For each table used in the query, any available selections are pushed to the leaves and
a search over imputation operations produces various possible plans, which are added to the cache. If no selections are available, a 
simple scan is added to the plan cache. Join order is jointly optimized with imputation placements and a set of
plans encompassing all necessary tables is produced. This set is now extended for any necessary grouping and aggregation
operations. After a similar step has been taken for projections, the planner now contains the plan frontier (i.e. the
best possible set from which to pick our final plan). The final step in the planner is to find all plans that are
$\alpha$-bound and return the one that has the lowest runtime: the $\alpha$-bound optimal plan.

@@

By restricting imputations to two classes: those that impute only attributes used in the operator and those that impute all attributes needed downstream in the query,
we limit the factor of plans at each operator to three: \textit{Drop}, \textit{Impute} over attributes used in operator, and \textit{Impute} over all attributes needed.
Recall that the planner maintains different plans for different dirty sets, keeping only those that are not dominated by other plans with the same dirty set. So in many
cases we can drop some of these intermediate plans generated at each operator. However, if the dirty sets tracked are distinct through the entire planning phase,
then we could explore $\mathcal{O}(3^{\# of operators}) * plans$.

>>>>

\begin{align*}
  \textsc{L}(q) &= \begin{cases}
     \{1\} \cup \textsc{L}(q') & q = \delta_C(q') \\
    \left\{\frac{1}{\sqrt{ \textsc{Attr}(q') * \textsc{Card}(q')}}\right\} \cup \textsc{L}(q')  & q = \mu_C(q') \\
    \textsc{L}(q_1') \cup \textsc{L}(q_2') & q = q_1' \bowtie_\psi q_2' \\
    \textsc{L}(q') & q = \sigma_\phi(q'), \\ & q = \pi_C(q') \\
    \emptyset & q = t\\
  \end{cases}\\
  \textsc{Loss}(q) &= \frac{\Sigma \textsc{L}(q)}{|\textsc{L}(q)|}
\end{align*}

\todo{Similar algorithm for computing Time?}

@@

The planner operates as follows.
First, it collects the set of attributes which are used by the operators in the plan or which are visible in the output of the plan (Lines~\ref{lst:line:attr-start}-\ref{lst:line:attr-end}).
This set will be used to determine which attributes are imputed.
Then, it constructs a plan cache. For each table used in the query, any available selections are pushed to the leaves and
a search over imputation operations produces various possible plans, which are added to the cache (Line~\ref{lst:line:sel}). If no selections are available, a 
simple scan is added to the plan cache (Line~\ref{lst:line:scan}). Join order is jointly optimized with imputation placements and a set of
plans encompassing all necessary tables is produced (Line~\ref{lst:line:join})\todo{Citation or explanation for join optimization algorithm}. This set is now extended for any necessary grouping and aggregation
operations (Line~\ref{lst:line:group}). After a similar step has been taken for projections, the planner now contains the plan frontier (i.e. the
best possible set from which to pick our final plan). The final step in the planner is to find all plans that are
$\alpha$-bound and return the one that has the lowest runtime: the $\alpha$-bound optimal plan.

@@

By restricting imputations to two classes: those that impute only attributes used in the operator and those that impute all attributes needed downstream in the query,
we limit the factor of plans at each operator to three: \textit{Drop}, \textit{Impute} over attributes used in operator, and \textit{Impute} over all attributes needed.
Recall that the planner maintains different plans for different dirty sets, keeping only those that are not dominated by other plans with the same dirty set. So in many
cases we can drop some of these intermediate plans generated at each operator. However, if the dirty sets tracked are distinct through the entire planning phase,
then we could explore $\mathcal{O}(3^{\# of operators}) * plans$\todo{this is unclear. what is plans?}.

### 2017-02-23 paper/algorithms/plan-toplevel.tex
paper: another partial pass over algorithm section

% \begin{algorithm}
\begin{algorithmic}[5]
  \Require{A set of tables $T$, a relation $F: T \times \Phi$ between tables and filter predicates, a relation $J : T \times \Psi \times T$ between tables and join predicates, a set of projection attributes $P$,  an optional set of grouping attributes $G$ and aggregator function $A$, and a parameter $\alpha \in [0, 1]$ that expresses trade-off performance and information loss}
  \Ensure{An optimized query plan.}

  \Function{Plan}{$T, F, J, P, G, A$}
  \State $C_g \gets \bigcup_{\phi \in F} \Call{Attr}{\phi}$ \Comment{Collect relevant attributes.} \label{lst:line:attr-start}
  \State $C_g \gets \bigcup_{\psi \in J} \Call{Attr}{\psi} \cup P \cup G \cup \Call{Attr}{A} \cup C_g$ \label{lst:line:attr-end}
  \State Let $Q$ be an empty plan cache.
  \For{$t \in T$} \Comment{Add selections to the plan cache.}
  \If {$\exists \phi: (t, \phi) \in F$}
  \State $Q[\{t\}] \lhd \Call{OptFilter}{Q, t, \phi, C_g}$
  \Else
  \State $Q[\{t\}] \lhd \{t\}$
  \EndIf
  \EndFor

  \State $\Call{OptJoin}{Q, T, J, C_g}$	\Comment{Optimize joins.}
  \State $B \gets Q[T]$ \Comment{Get the best plans for all tables.}
  \If{$G \neq \emptyset$} \Comment{Add optional group \& aggregate.}
  \State $B \gets \bigcup_{q \in B} \Call{OptGroupBy}{q, G, A, C_g}$
  \EndIf

  \State $B \gets \bigcup_{q \in B} \Call{OptProject}{q, P}$ \Comment{Add projections.}
  \State \Return $p \in B$ s.t. $p$ is $\alpha$-bound optimal  \Comment{Return fastest plan satisfying $\alpha$ bound.}
  \EndFunction

>>>>

% \begin{algorithm}
\begin{algorithmic}[1]
  \Require{A set of tables $T$, a relation $F: T \times \Phi$ between tables and filter predicates, a relation $J : T \times \Psi \times T$ between tables and join predicates, a set of projection attributes $P$,  an optional set of grouping attributes $G$ and aggregator function $A$, and a parameter $\alpha \in [0, 1]$ that expresses trade-off performance and information loss}
  \Ensure{An optimized query plan.}

  \Function{Plan}{$T, F, J, P, G, A$}
  \State $C_g \gets \bigcup_{\phi \in F} \Call{Attr}{\phi}$ \Comment{Collect relevant attributes.}~\label{lst:line:attr-start}
  \State $C_g \gets \bigcup_{\psi \in J} \Call{Attr}{\psi} \cup P \cup G \cup \Call{Attr}{A} \cup C_g$ \label{lst:line:attr-end}
  \State Let $Q$ be an empty plan cache.
  \For{$t \in T$} \Comment{Add selections to the plan cache.}
  \If {$\exists \phi: (t, \phi) \in F$}
  \State $Q[\{t\}] \lhd \Call{OptFilter}{Q, t, \phi, C_g}$~\label{lst:line:sel}
  \Else
  \State $Q[\{t\}] \lhd \{t\}$~\label{lst:line:scan}
  \EndIf
  \EndFor

  \State $\Call{OptJoin}{Q, T, J, C_g}$	\Comment{Optimize joins.}~\label{lst:line:join}
  \State $B \gets Q[T]$ \Comment{Get the best plans for all tables.}
  \If{$G \neq \emptyset$} \Comment{Add optional group \& aggregate.}~\label{lst:line:group}
  \State $B \gets \bigcup_{q \in B} \Call{OptGroupBy}{q, G, A, C_g}$
  \EndIf

  \State $B \gets \bigcup_{q \in B} \Call{OptProject}{q, P}$ \Comment{Add projections.}
  \State \Return $p \in B$ s.t. $p$ is $\alpha$-bound optimal.
  \EndFunction
